Bayesian Deep Gaussian Processes for Correlated Functional Data: A Case Study in Cosmological Power Spectra

### 2024-10-16 paper/main.tex
revise abstract

\begin{abstract} 
Understanding the structure of our universe and the distribution of matter is an 
area of active research.  As cosmological surveys grow in complexity, developing 
surrogate models or emulators to efficiently predict matter power spectra is 
important.  We are particularly motivated by the Mira Titan Universe simulation
suite which, for a specified cosmology parameterization, provides multiple response
curves of various fidelities, including correlated functional realizations.
In this work, we propose a Bayesian deep Gaussian process (DGP) hierarchical 
model to synthesize all the information to estimate the underlying matter power
spectra, while providing effective uncertainty quantification.  This model
extends previous works on Bayesian DGPs from scaler responses to functional ones.
We deploy basis function representations of the estimated spectra to train another
surrogate for the prediction of unobserved cosmologies.  Our method performs well
compared to the benchmark cosmological emulator (Cosmic Emu).  We provide
reproducible code in a public git repository.
\end{abstract}

>>>>

\begin{abstract} 
Understanding the structure of our universe and the distribution of matter is an 
area of active research.  As cosmological surveys grow in complexity, the development 
of emulators to efficiently and effectively predict matter power spectra is essential.  
We are particularly motivated by the Mira Titan Universe simulation
suite which, for a specified cosmological parameterization (termed a ``cosmology''), 
provides multiple response curves of various fidelities, including correlated 
functional realizations.  Our objective is two-fold.  First, we seek to 
estimate the underlying true matter power spectra, with appropriate uncertainty 
quantification (UQ), from all of the provided curves.  To this end, we propose a 
novel Bayesian deep Gaussian process (DGP) hierarchical model which synthesize 
all the information and estimates the underlying matter power spectra, while providing 
effective UQ.  Our model extends previous work on Bayesian DGPs from scaler responses 
to functional ones.  Second, we aim to leverage simulation results of various 
cosmologies in order to accurately predict the entire matter power spectra for an 
unobserved cosmology.  For this we leverage basis function representations 
of the estimated spectra to train a separate Gaussian process emulator.  
Our method performs well in synthetic exercises and against the benchmark cosmological 
emulator (Cosmic Emu).  We provide reproducible code in a public git repository.
\end{abstract}

\noindent \textbf{Keywords:} surrogate, computer experiment, uncertainty quantification,
principal components analysis, ???
